---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AMMIT

<!-- badges: start -->
<!-- badges: end -->

AMMIT is designed to streamline the processing of many mIHC samples, by applying a skew-normal mixture model trained on a small subset of samples.

## Installation

You can install the development version of AMMIT from [GitHub](https://github.com/) with:

``` r
library(devtools)
devtools::install_github("neeson-lab/AMMIT")
```

## Basic AMMIT workflow 

### Reading in your data

You can read in your data from a HALO file (often titled "Total_Object_Results.csv") or you can feed it in directly. Data is converted into the SpatialExperiment class. 

The data for the following examples is accessible within the AMMIT package:

``` {r, include=F}
library(AMMIT)
library(SpatialExperiment)
``` 
  
```{r}
data("ammit_spe_dummy")
eg_markers <-  c("M1", "M2", "M3", "M4", "M5", "M6", "M7")
```

The object contains two 'assays' - a 'data' assay which contains the intensity data for each marker and cell, and a 'reference' assay which contains the manually-thresholded state for each marker and cell. Note that in instances where you are not training or validating, your object does not need to have reference data. 

As we progress with AMMIT, most of our results will be stored in the `rowData` of this object, which carries metadata for each individual marker (the rows of our SpatialExperiment object), independent of cells. Later down the track, cell phenotypes and so on will be stored in the `colData`. 

It was built like this: 

```{r}
# ammit_spe_dummy <- build_ammit_spe(method = "HALO",
#                                    markers = eg_markers,
#                                    reference = TRUE,
#                                    HALO_filepath = system.file("extdata", "halo_output.csv", package="AMMIT"),
#                                    locations = c("Nucleus", "Cytoplasm", "Nucleus", "Cytoplasm", "Cytoplasm", "Cytoplasm", "Cytoplasm"),
#                                    filter_dapi = TRUE)
```

Note that if your data does not include DAPI positivity classification (e.g. segmentation was completed without setting a threshold for DAPI positivity) you can set `filter_dapi` to `FALSE`. 

The reference parameter here is set to `TRUE`, indicating this data has already been manually-thresholded and can be used to train (or validate) the strictness parameter for the AMMIT workflow. However, to do this we also need the manual thresholds themselves. While it is possible to read them in when building the SpatialExperiment object, here we will apply a simple function to pull them from the reference and intensity data for each marker. 

```{r}
ammit_spe_dummy <- infer_manual_thresholds(ammit_spe_dummy, eg_markers)
```

### Plotting marker intensity densities

We might like to examine our data at this stage. AMMIT allows for the basic plotting of the distribution of the marker intensities using the `plot_intensities` function. Let's take a look at the marker M7:

```{r}
plot_intensities(ammit_spe_dummy,
                 markers = "M7",
                 thresholds = "manual",
                 transformation = "asinh",
                 censor_zeroes = TRUE)
```

### Unmixing: the basis of AMMIT

AMMIT is built around the unmixing of your marker intensity distributions into several component distributions. We apply a mixture of skew-normal distributions model to pull out `k` components. Generally, we use a `k` of 2 or 3, which accounts for a distribution of 'negative', 'positive' and (in the case of `k` = 3) 'background' intensity values. 

By default, we censor values of zero and transform the data with an arcsinh transformation prior to unmixing.

When unmixed, parameters for the component distributions are stored in the rowData of the SpatialExperiment, as below:

```{r}
ammit_spe_dummy <- unmix_intensities(ammit_spe_dummy, markers="M7", k=2)
SummarizedExperiment::rowData(ammit_spe_dummy)["M7", ]
```

We can then plot the results of our unmixing with `plot_mixture`:

```{r}
plot_mixture(ammit_spe_dummy, markers="M7", thresholds="manual", transformation="asinh")
```
  
### Training: inferring the strictness parameter

Before running AMMIT on your full cohort, you'll first need to train the strictness parameter to be applied for each marker. This 
